### 2015-01-08 inst/doc/sparseHessianFD.Rmd
it works, with lots of new documentation

---
title:  Using sparseHessianFD
author:  Michael Braun
date:  "`r Sys.Date()`"
output:
  html_vignette:
    toc:  false
  pdf_vignette:
    toc:  false	
header-includes:
  \usepackage{amsmath}
vignette: >
  %\VignetteIndexEntry{sparseHessian example}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

The **sparseHessianFD** package provides an interface to ACM TOMS Algorithm
636 for estimating a Hessian that is sparse, in that a large proportion of
the cross-partial derivatives are zero.  The user provides the following:

1.  an R function that returns the value of the objective function
whene valuated at a variable vector $P$;
2.  an R function that returns the gradient of that objective function
at $P$; and
3.  the row and column indices for the non-zero elements in the lower triangle of the Hessian.

The non-zero elements are estimated through finite differencing of
the gradients in a way that exploits the sparsity pattern of the
Hessian.

There are, of course, other methods to estimate sparse Hessians, such
as automatic differentiation (AD).  However, operator-overloading
implementations of AD require the objective function to be written
using specialized libraries.  These libraries are available for C++,
Fortran, Matlab, Python, and possibly other programming languages and
environments.  At the present, there are no AD implementations that
are "native" to R.  Other than deriving the Hessian analytically,
finite differencing remains the best option when the objective and
gradient functions are written purely in R.

## Example function

Before going into the details of how to use the package, let's
consider the following example of an objective function with a sparse Hessian.
 Suppose we have a dataset of $N$ households, each with $T$ opportunities to purchase a particular product.  Let $y_i$ be the number of times household $i$ purchases the product, out of the $T$ purchase opportunities.  Furthermore, let $p_i$ be the probability of purchase; $p_i$ is the same for all $T$ opportunities, so we can treat $y_i$ as a binomial random variable.  The purchase probability $p_i$ is heterogeneous, and depends on both $k$ continuous covariates $x_i$, and a heterogeneous coefficient vector $\beta_i$, such that
$$
  p_i=\frac{\exp(x_i'\beta_i)}{1+\exp(x_i'\beta_i)},~i=1 ... N
$$

The coefficients can be thought of as sensitivities to the covariates, and they are distributed across the population of households following a multivariate normal distribution with mean $\mu$ and covariance $\Sigma$.   We assume that we know $\Sigma$, but we do not know $\mu$.  Instead, we place a multivariate normal prior on $\mu$, with mean $0$ and covariance $\Omega_0$.  Thus, each $\beta_i$, and $\mu$ are $k-$dimensional vectors, and the total number of unknown variables in the model is $(N+1)k$.

The log posterior density, ignoring any normalization constants, is
$$
  \log \pi(\beta_{1:N},\mu|Y, X, \Sigma_0,\Omega_0)=\sum_{i=1}^Np_i^{y_i}(1-p_i)^{T-y_i}
  -\frac{1}{2}\left(\beta_i-\mu\right)'\Sigma^{-1}\left(\beta_i-\mu\right)
-\frac{1}{2}\mu'\Omega_0^{-1}\mu
$$

```{r, echo=FALSE}
N <- 6
k <- 2
nv1 <- (N+1)*k
nels1 <- nv1^2
nnz1 <- (N+1)*k^2 + 2*N*k^2
nnz1LT <- (N+1)*k*(k+1)/2 + N*k^2
Q <- 1000
nv2 <- (Q+1)*k
nels2 <- nv2^2
nnz2 <- (Q+1)*k^2 + 2*Q*k^2
nnz2LT <- (Q+1)*k*(k+1)/2 + Q*k^2
options(scipen=999)
```

Since the $\beta_i$ are drawn iid from a multivariate normal,
$\dfrac{\partial^2\log\pi }{\partial\beta_i\beta_j}=0$ for all $i\neq
j$.  We also know that all of the $\beta_i$ are correlated with
$\mu$.  The structure of the Hessian depends on how the variables are
ordered within the vector. One such ordering is to group all of the
coefficients for each unit together.

$$
\beta_{11},...,\beta_{1k},\beta_{21},...,\beta_{2k},...~,~...~,~\beta_{N1}~,~...~,~\beta_{Nk},\mu_1,...,\mu_k
$$

In this case, the Hessian has a "block-arrow" structure.  For example,
if $N=`r N`$ and $k=`r k`$, then there are `r nv1` total variables, and the Hessian will have the following pattern.

```{r, echo=FALSE}
M <- as(kronecker(diag(N),matrix(1,k,k)),"lMatrix")
M <- rBind(M, Matrix(TRUE,k,N*k))
M <- cBind(M, Matrix(TRUE, k*(N+1), k))
print(M)
```

There are `r nels1` elements in this symmetric matrix, but only  `r nnz1` are
non-zero, and only `r nnz1LT` values are unique.  Although the reduction in
RAM from using a sparse matrix structure for the Hessian may be
modest, consider what would happen if $N=`r Q`$ instead.  In that case,
there are `r nv2` variables in the problem, and more than $`r 
floor(nels2/10^6)`$ million
elements in the Hessian.  However, only $`r nnz2`$ of those elements are
non-zero.  If we work with only the lower triangle of the Hessian we only need to work with
only `r nnz2LT` values.

Another possibility is to group coefficients for each covariate
together.

$$
\beta_{11},...,\beta_{1N},\beta_{21},...,\beta_{2N},...,...,\beta_{k1},...,\beta_{kN},\mu_1,...,\mu_k
$$

Now the Hessian has an "off-diagonal" sparsity pattern.

```{r, echo=FALSE}
M <- as(kronecker(matrix(1,k,k), diag(N)),"lMatrix")
M <- rBind(M, Matrix(TRUE,k,N*k))
M <- cBind(M, Matrix(TRUE, k*(N+1), k))
print(M)
```

In both cases, the number of non-zeros is the same.  The
block-diagonal case may lead to slightly faster initialization,
but repeated computation of the Hessian will take the same time as
with the "off-diagonal" case.

## Using the package

The functions for computing the objective function, gradient and
Hessian for this example are in the R/ex\_funcs.R file.  The package
also includes a sample dataset with simulated data from the binary choice example.

To start, we load the data, set some dimension parameters, set prior
values for $\Sigma^{-1}$ and $\Omega^{-1}$, and simulate a
vector of variables at which to evaluate the function.

```{r}
set.seed(123)
data(binary)
Y <- binary$Y
X <- binary$X
T <- binary$T
N <- length(Y)
k <- NROW(X)
nvars <- as.integer(N*k + k)
P <- rnorm(nvars) ## random starting values
inv.Sigma <- rWishart(1,k+5,diag(k))[,,1]
inv.Omega <- diag(k)
```

```r
options(scipen=-999)
```

This dataset represents the simulated choices for $N= `r N`$ customers
over $T= `r T`$ purchase opportunties, where the probability of purchase
is influenced by $k= `r k`$ covariates.

The objective function for the binary choice example is `binary.f` and the gradient function is
`binary.grad`.  The first argument to both is the variable vector, and
the argument lists must be the same for both.  For this example, we
need to provide the data ($X$ and $Y$) and the prior parameters
($\Sigma^{-1}$ and $\Omega^{-1}$). The functions also have an
`order.row` argument to change the ordering of the variables (and
thus, the sparsity pattern). If `order.row=TRUE`, then the Hessian
will have an off-diagonal pattern.  If `order.row=FALSE`, then the
Hessian will have a block-arrow pattern.

For testing and demonstration purposes, we also have a `binary.hess`
function that returns the Hessian as a sparse `dgCMatrix` object (see
the Matrix package).

```{r, tidy=TRUE}
true.f <- binary.f(P, Y=Y, X=X, T=T, inv.Sigma=inv.Sigma, inv.Omega=inv.Omega, order.row=FALSE)
true.grad <- binary.grad(P, Y=Y, X=X, T=T, inv.Sigma=inv.Sigma, inv.Omega=inv.Omega, order.row=FALSE)
true.hess <- binary.hess(P, Y=Y, X=X, T=T, inv.Sigma=inv.Sigma, inv.Omega=inv.Omega, order.row=FALSE)
```

The sparsity pattern of the Hessian is specified by two **integer**
vectors:  one each for the row and column indices of the non-zero
elements of the ***lower triangular part*** of the Hessian.  If you
happen have have an example of a matrix with the same sparsity pattern
of the Hessian you are trying to compute, you can use the following
convenience function to extract the appropriate index vectors.

```{r}
pattern <- Matrix.to.Coord(tril(true.hess))
str(pattern)	    	    
```

Next, we create a new instance of a sparseHessianFD object with an
"initial variable" $P$, and the row and column indices of the non-zero
elements in the lower triangle of the Hessian.  We also pass in any
other arguments for `binary.f` and `binary.grad`.  We accept the default
values for other arguments to `sparseHessianFD.new`.
```{r}
obj <- sparseHessianFD.new(P, binary.f, binary.grad,
       rows=pattern$iRow, cols=pattern$jCol,
       X=X, Y=Y, T=T, inv.Sigma=inv.Sigma, inv.Omega=inv.Omega,
       order.row=FALSE)
```

Now we can evaluate the function value, gradient and Hessian through
`obj`.

```{r}
f <- obj$fn(P)
gr <- obj$gr(P)
hs <- obj$hessian(P)
```
Note that the member functions in the sparseHessianFD class take only
one argument:  the variable vector.  All of the other arguments are
stored in `obj`.

Do we get the same results that we would get after calling `binary.f`,
`binary.grad` and `binary.hess` directly?  Let's see.
```{r}

all.equal(f, true.f)
all.equal(gr, true.grad)
all.equal(hs, true.hess)	      	     	     
```

If there is any difference, keep in mind that `hs` is a numerical
estimate that is not always exact.  I certainly wouldn't worry about
mean relative differences smaller than, say, $10^{-6}$.

### Speed comparison

Instead of using this package, we could treat the Hessian as dense,
and use the hessian function numDeriv package. The advantage of the
numDeriv package is that it does not require the gradient.  However,
you can see that it takes some time to run.

```{r, tidy=TRUE}
library(numDeriv)
hess.time <- system.time(H1 <- obj$hessian(P))	  
fd.time <- system.time(H2 <- hessian(obj$fn, P))
H2 <- drop0(H2, 1e-7) ## treat values < 1e-8 as zero
print(hess.time)
print(fd.time)	   
```

The sparseHessianFD package can be substantially faster than estimating a dense
Hessian by brute force finite differencing.  The cost of this speed up
is that the user does need to provide the gradient and the sparsity
structure.  As with everything in life, there are trade-offs.